Collision rates of planetesimals near mean-motion resonances

### 2021-02-23 main.tex
Remove boldface edit marks

\begin{abstract}
\textbf{In circumstellar discs, collisional grinding of planetesimals produces second-generation dust. While it remains unclear whether this ever becomes a major component of the total dust content, the presence of such dust, and potentially the substructure within, it can be used to explore a disc's physical conditions. A perturbing planet produces nonaxisymmetric structures and gaps in the dust, regardless of its origin. The dynamics of planetesimals, however, will be very different than that of small dust grains due to weaker gas interactions. Therefore, planetesimal collisions could create dusty disc structures that would not exist otherwise. In this work, we use N-body simulations to investigate the collision rate profile of planetesimals near mean-motion resonances. We find that a distinct bump or dip feature is produced in the collision profile, the presence of which depends on the libration width of the resonance and the separation between the peri- and apocenter distances of the edges of the resonance. The presence of one of these two features depends on the mass and eccentricity of the planet. Assuming that the radial dust emission traces the planetesimal collision profile, the presence of a bump or dip feature in the dust emission at the 2:1 mean-motion resonance can constrain the orbital properties of the perturbing planet. This assumption is valid, so long as radial drift does not play a significant role during the collisional cascade process. Under this assumption, these features in the dust emission should be marginally observable in nearby protoplanetary disks with ALMA.}
\end{abstract}

@@

Recent observations of circumstellar discs by ALMA have revealed a rich variety of substructure. Features such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the discs. In many gas-rich systems, these gap features are argued to indicate the 
presence of a giant planet, either embedded in the disc \citep{2015MNRAS.453L..73D} or orbiting \textbf{adjacent to it}, although the details of gas-dust 
interactions can have strong effects on the resulting morphologies
\citep{2018ApJ...866..110D}. In gas-poor/debris systems, a giant
planet perturber can also influence 
the structure of the dust continuum emission. For example, a misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}, highly eccentric perturbers can produce structures through secular interactions 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}, and mean motion resonances (MMRs) can open gaps
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

@@

\citet{2013ApJ...777L..31D} showed that collisional dust that is generated near the gap opened by a giant planet in \textbf{the inner regions of} a \textbf{transition} disc should 
produce a distinct observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological 
differences in the dust emission could be used to determine whether the planet had a circular or eccentric orbit \citep{2016ApJ...820...29D}. Moreover, 
substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues that could be used to constrain the properties of 
the planet.  The width of a resonance is set by both the mass of the perturbing planet and the unperturbed eccentricity of the planetesimals, 
which is in turn set by the eccentricity of the planet through secular forcing. Therefore, \textbf{the planetesimal collision profile and potentially} the structure of second generation 
dust produced near MMRs may encode information about both the mass and eccentricity of the planet.

The dynamics governing the motion of bodies near MMRs is extremely nonlinear, as is determining what the collision rates between planetesimals 
should look like in these regions. For a collection of bodies massive enough to experience the effects of gravitational focusing, a large eccentricity 
dispersion tends to reduce the probability of collision, while enhancements in surface density tends to increase it. Due to conservation of the Jacobi 
energy, MMRs simultaneously enhance the local eccentricity dispersion and also enhance the surface density adjacent to the resonance 
\citep{2000Icar..143...45R, 2017ApJ...850..103B}. Unfortunately, collision detection in an N-body simulation is extremely computationally expensive. 
So far, studies of planetesimal dynamics near MMRs \textbf{in which the planetesimals are directly resolved} have involved either collisionless test particles 
\citep{2017ApJ...850..103B, 2016ApJ...818..159T, 2018ApJ...857....3T} or \textbf{integration times that do not fully capture the dynamics of the resonances} 
\citep{2000Icar..143...45R, 2013ApJ...777L..31D}.

To further elucidate this subject, we use the tree based N-body code {\sc ChaNGa}\citep{2008IEEEpds...ChaNGa, 2015AphCom..2..1}  to follow the 
collisional evolution of a planetesimal disc under the gravitational influence of a Jupiter-sized body. Because particle positions are sorted into a tree 
structure, neighbor finding and collision detection can be done quickly and efficiently. This considerably relaxes the constraints on resolution and 
integration time. With this toolset, we explore the collision rate structure of a planetesimal disc in the vicinity of mean-motion resonances with a planet. In 
particular, we would like to determine (1) \textbf{what dynamics govern the collision rate profile near MMRs, (2)} whether MMRs leave a detectable signature in the
collisionally-generated dust, and \textbf{(3)} whether these signatures can be used to determine or constrain the orbital properties of the perturbing planet.

This work is organized in the following way: In section \ref{sec:dynamics}, we provide an overview of the relevant dynamics that drive the evolution 
of a planetesimal disc under the gravitational influence of an external perturber. In section \ref{sec:sims} we provide an overview of the N-body code 
used and describe the initial conditions chosen for five simulations in which a perturbing giant planet is given various masses  and eccentricities. Section
\ref{sec:results} presents the results of these simulations \textbf{and we take an in-depth look at the collision rate profiles of the planetesimals near the MMRs. 
In section \ref{sec:dust},  we extrapolate the collision rate down to smaller bodies and motivate a direct correspondence between the planetesimal collision profile
and the resulting dust profile. In section \ref{sec:constrain}, w}e use the resolved collisions to generate synthetic dust 
emission profiles that would be detected with observing facilities like ALMA. \textbf{Under our simplifying assumptions, we show} that a characteristic bump 
or dip feature appears in the dust emission near the interior 2:1 MMR, the presence of which depends on the mass and eccentricity of the perturbing 
planet. \textbf{If the 2:1 MMR can be identified (presumably, by identifying another prominent MMR, such as the 3:1, and measuring the spacing between the two),} we discuss the 
potential for using this feature to place constraints on the mass and eccentricity of the planet\textbf{. In addition, we highlight the caveats of connecting the dust and planetesimal collision 
profiles in such a simple way.} \textbf{Finally, we} conclude in section \ref{sec:conclusions}.

@@

\textbf{We begin by providing a description of the dynamical effects responsible for shaping the orbital distribution of the planetesimals. The purpose of this is twofold: (1) to motivate
the initial conditions used for the simulations described in section \ref{sec:ics} and (2) to justify the exclusion of certain physical effects from our simulations. Here, we focus solely on the 
physics relevant for full-sized ($\sim$100 km) planeteismals and save a discussion of the effects on smaller bodies generated through collisions for section \ref{sec:dust}.}

@@

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and a planetesimal, 
respectively. $e_{g}$ and $\varpi_{g}$ are the eccentricity and
longitude of pericenter of the giant planet, and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{1999ssd..book.....M} \textbf{(ch. 6, pg. 237, eq. 6.67)} as

@@

\textbf{For the case of a Jupiter mass planet at 5.2 au perturbing a test particle at 3 au, the timescale for secular forcing is approximately 12,000 years. The details of this calculation
can be found in appendix \ref{sec:sec_forcing_timescale}.}

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a
body, with $M$ being the mean anomaly. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Furthermore, variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citep{1999ssd..book.....M}. An important 
point to note, which we will revisit later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of $\phi$. 
\textbf{Calculations for the widths of first and second order interior MMRs are shown in appendix \ref{sec:libration}, along with a timescale for the libration. For the most prominent interior 
mean-motion resonances, including the 2:1 and 3:1, the timescale associated with these oscillations driven by a Jupiter mass planet is  $\sim$ 1,000 - 2,000 years.}

@@

\textbf{A simple analytic model for the collision rate of a planetesimal population is given by \citet{1969edo..book.....S} as}

@@

\noindent\textbf{where $n$ is the number density of the population, $s$ and $m$ are the radii and masses of the bodies and $v$ is their typical encounter velocity. The encounter velocity is often described in terms of the rms eccentricity $\left<e^{2}\right>^{1/2}$ and inclination $\left<i^{2}\right>^{1/2}$ of the population (e.g. \citet{1993ARA&A..31..129L}) as}

@@

\noindent\textbf{where $v_{k}$ is the local Keplerian velocity. The second term in equation \ref{eq:saf} can be thought of as an additional enhancement to the collision cross section due to 
gravitational focusing. When the typical encounter velocity is small compared to the mutual escape velocity of the planetesimals, the collision cross section greatly exceeds the geometric 
value. An important feature of equation \ref{eq:saf} is that for a fixed value of $n$, the collision rate exhibits a global minimum as a function of $v$. For small $v$, gravitational focusing 
facilitates more collisions, while for large $v$ the encounter rate simply becomes so great that the collision rate again increases, even though gravitational focusing is mostly surpressed. 
This point will become relevant in section \ref{sec:vary_ecc} when we examine the qualitative changes in the collision rate near mean-motion resonances.}

\textbf{Although equation \ref{eq:saf} works well to describe the collision rate for a homogeneous collection of planetesimals, some problems arise when regions of commensurabilities are 
introduced. Namely, the resonances cause the orbits of planetesimals to precess, and an interface between secularly aligned and randomly oriented orbits arise on each side of the 
resonance. Additionally, an interface between dynamically cold and hot planetesimals develop. These two effects cause the number density, collision cross section and encounter velocity 
to rather abruptly vary across the boundaries of the resonance. For this reason, an N-body treatment in which collisions are directly resolved is necessary to understand how the collision 
rate varies near the MMRs.}

@@

\textbf{Over the course of many orbits, the residual gas from the primordial nebula can damp the eccentricities and inclinations of bodies. For a planetesimal-sized body, gas drag operates 
in the Stokes regime and the timescale for aerodynamic forces to significantly alter its relative velocity is given by \citep{1976PThPh..56.1756A}}

@@

\noindent \textbf{where $m$ and $s$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a
density of $4 \times 10^{-10}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5,000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km 
body with a density of 2 g cm$^{-3}$ \textbf{(assuming a mixture of ice and rock)} the stopping timescale is on the order of 1 Myr. This is much longer than the timescales associated with 
secular forcing and libration due to mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially shorter the planetesimal 
formation timescale. For this reason, we do not model the effects of gas drag in the simulations, although we consider its effects when constructing initial conditions, which is discussed in 
the next section.}

@@

To follow the dynamical and collisional evolution of a planetesimal disc, we use the highly parallel N-body code {\sc ChaNGa} 
\footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}}. 
This code, which is written in the {\sc CHARM++} parallel programming language, was originally designed for cosmology simulations 
and has been shown to perform well on up to half a million processors \citep{2015AphCom..2..1}. \textbf{Using similar methods to {\sc PKDGRAV}, which has been used for numerous 
studies of planet formation \citep{2000Icar..143...45R, 2005ApJ...625..427L, 2013ApJ...777L..31D, 2015ApJ...806...23L}}, {\sc ChaNGa} calculates 
gravitational forces using a modified Barnes-Hut \citep{1986Natur.324..446B} tree with hexadecapole expansions of the moments \textbf{and integrates the equations of motion using a 
kick-drift-kick leapfrog scheme}. All of the simulations we perform use a node opening criterion of $\theta_{BH}$ = 0.7. More information about the code can be found in 
\citet{2008IEEEpds...ChaNGa}.

\textbf{We have recently modified {\sc ChaNGa} to handle solid-body collisions between particles, by assigning them a fixed radius, rather than treating them as tracers of a fluid with a 
characteristic softening length. We provide a brief summary of the collision model here, although a full description is provided in \citet{2019MNRAS.489.2159W}. This work is largely based 
on the solid-body collision implementation in {\sc PKDGRAV}, which is detailed in \citet{1994MNRAS.269..493R} and \citet{2000Icar..143...45R}. Imminent collisions are found during the 
drift phase of each time step by extrapolating the positions of bodies forward using the velocity calculated during the first kick. For each body, the nearest 64 neighbors are tested for an 
imminent collision. If an imminent collision is detected, the two particles of mass $m_{1}$ and $m_{2}$ are merged together together to form one single larger body with the same density 
and a mass of $m_{1} + m_{2}$. The resulting body is then imparted
with the center of mass position and velocity of the two
colliders. Because the resolution of a collision can result in 
another imminent collision, these events must be handled one at a time, with the soonest collision being resolved first. For this reason, another collision search is run each time a collision 
is resolved, which continues until there are no more imminent collisions during the current time step. During the course of the simulation, no debris is created by collisions. To model the 
collisionally generated dust distribution, we use the statistics of
the resolved planetesimal collisions to build a dust profile. This process is discussed in detail in section \ref{sec:dust}.}

@@

In total, five simulations are run, which are listed in table \ref{tab:sims}. The second is a ``nominal'' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass \textbf{or eccentricity is altered} by a factor of two \textbf{from} the nominal value. In all cases, the perturbing giant is placed on a 5.2 au orbit around a 1 
$M_{\odot}$ star. The planetesimal disc extends from 2.2 to 3.8 au, which covers the two most prominent mean-motion resonances with the giant planet, the 2:1 at 3.27 au and the 3:1 at 
2.5 au. The planetesimal disc follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$ \textbf{, which corresponds to a mixture of ice and rock,} and 
\textbf{are given a} diameter of 300 km. \textbf{This choice of parameters is meant to mimic the setup used in simulation B from \citet{2000Icar..143...45R}. One difference from the 
aformentioned study is that we use a narrower annulus, which reduces the number of particles required from $10^6$ down to roughly 500,000. This allows us to use a finer base timestep 
size, which, as we will discuss in section \ref{sec:results}, appears to be the reason why no features were seen in the collision profile near the resonances by \citet{2000Icar..143...45R}}.

\textbf{Another difference from \citet{2000Icar..143...45R} is that} the dynamical effects of secular forcing by the giant planet, along with the effects of viscous stirring and gas drag on the 
planetesimals, are built into the initial conditions. \textbf{Although we do not model the effects of gas drag on the planetesimals during the simulation, we construct the initial conditions 
such that the effects of gas drag and viscous stirring are in balance. The viscous stirring timescale of the planetesimal disc is much longer than our chosen integration time, and so the 
dynamical excitation of the disc (excluding resonances) stays constant, even without the inclusion of damping forces from the gas.} This is done by first calculating the equilibrium 
eccentricity $e_{eq}$ due to viscous stirring and gas drag as a function of semimajor axis according to equation 12 of \citet{2002ApJ...581..666K}. The eccentricities of the bodies are 
drawn randomly from a Rayleigh distribution with a mode of $e_{eq}$, while the inclinations are drawn from a similar distribution with a mode of $e_{eq}$/2 \citep{1993MNRAS.263..875I}. 
The arguments of perihelion $\omega$, longitude of ascending nodes $\Omega$ and the mean anomalies $M$ of the bodies are drawn uniformly $\in [0, 2 \pi)$.

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. \textbf{Through experimentation}, we have found that using a base timestep size with {\sc ChaNGa} of 3\% of an orbital dynamical time
at the inner edge of the disc keeps the integration symplectic. Although doing so preserves orbital frequencies, the errors associated with
precession frequencies do not average out to zero. This is especially important given the effects of secular forcing by the planet. To mitigate
this, we further reduce the base timestep size by a factor of 4 and use $\Delta t$ = 0.0025 yr. Doing so prevents the longitude of perihelia of
planetesimals at the inner edge of the disc, where this artificial precession is most severe, from drifting by more than the intrinsic spread in $\varpi$ due to the
free eccentricity over the course of our integrations.

An additional timescale is set by the dynamical time of the planetesimals ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. \textbf{This timescale must be resolved in order to properly 
follow close gravitational encounters between these bodies}. To resolve the base time step and the dynamical timescale of planetesimals simultaneously, we use a two-tiered time 
stepping scheme, following\citep{2015ApJ...806...23L}. To start, all bodies are placed on the base time step. A first pass of collision detection is then run in which the radii of all bodies are 
inflated by a factor of 2.5. Any bodies with imminent collisions predicted using the inflated radii are placed on a time step that is a factor of 16 smaller than the orbital time step. Although 
this is still a factor of roughly 7 larger than the dynamical time of a planetesimal, we found no difference in the collision rate when using any smaller of a minimum time step size.The 
purpose of the two-tiered scheme is to properly resolve the gravitational interactions between any bodies that undergo a close encounter. This prevents the coarser base time step from 
reducing the effectiveness of gravitational focusing, while minimizing the additional computational expense.

>>>>

\begin{abstract}
In circumstellar discs, collisional grinding of planetesimals produces second-generation dust. While it remains unclear whether this ever becomes a major component of the total dust content, the presence of such dust, and potentially the substructure within, it can be used to explore a disc's physical conditions. A perturbing planet produces nonaxisymmetric structures and gaps in the dust, regardless of its origin. The dynamics of planetesimals, however, will be very different than that of small dust grains due to weaker gas interactions. Therefore, planetesimal collisions could create dusty disc structures that would not exist otherwise. In this work, we use N-body simulations to investigate the collision rate profile of planetesimals near mean-motion resonances. We find that a distinct bump or dip feature is produced in the collision profile, the presence of which depends on the libration width of the resonance and the separation between the peri- and apocenter distances of the edges of the resonance. The presence of one of these two features depends on the mass and eccentricity of the planet. Assuming that the radial dust emission traces the planetesimal collision profile, the presence of a bump or dip feature in the dust emission at the 2:1 mean-motion resonance can constrain the orbital properties of the perturbing planet. This assumption is valid, so long as radial drift does not play a significant role during the collisional cascade process. Under this assumption, these features in the dust emission should be marginally observable in nearby protoplanetary disks with ALMA.
\end{abstract}

@@

Recent observations of circumstellar discs by ALMA have revealed a rich variety of substructure. Features such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the discs. In many gas-rich systems, these gap features are argued to indicate the 
presence of a giant planet, either embedded in the disc \citep{2015MNRAS.453L..73D} or orbiting adjacent to it, although the details of gas-dust 
interactions can have strong effects on the resulting morphologies
\citep{2018ApJ...866..110D}. In gas-poor/debris systems, a giant
planet perturber can also influence 
the structure of the dust continuum emission. For example, a misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}, highly eccentric perturbers can produce structures through secular interactions 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}, and mean motion resonances (MMRs) can open gaps
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

@@

\citet{2013ApJ...777L..31D} showed that collisional dust that is generated near the gap opened by a giant planet in the inner regions of a transition disc should 
produce a distinct observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological 
differences in the dust emission could be used to determine whether the planet had a circular or eccentric orbit \citep{2016ApJ...820...29D}. Moreover, 
substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues that could be used to constrain the properties of 
the planet.  The width of a resonance is set by both the mass of the perturbing planet and the unperturbed eccentricity of the planetesimals, 
which is in turn set by the eccentricity of the planet through secular forcing. Therefore, the planetesimal collision profile and potentially the structure of second generation 
dust produced near MMRs may encode information about both the mass and eccentricity of the planet.

The dynamics governing the motion of bodies near MMRs is extremely nonlinear, as is determining what the collision rates between planetesimals 
should look like in these regions. For a collection of bodies massive enough to experience the effects of gravitational focusing, a large eccentricity 
dispersion tends to reduce the probability of collision, while enhancements in surface density tends to increase it. Due to conservation of the Jacobi 
energy, MMRs simultaneously enhance the local eccentricity dispersion and also enhance the surface density adjacent to the resonance 
\citep{2000Icar..143...45R, 2017ApJ...850..103B}. Unfortunately, collision detection in an N-body simulation is extremely computationally expensive. 
So far, studies of planetesimal dynamics near MMRs in which the planetesimals are directly resolved have involved either collisionless test particles 
\citep{2017ApJ...850..103B, 2016ApJ...818..159T, 2018ApJ...857....3T} or integration times that do not fully capture the dynamics of the resonances 
\citep{2000Icar..143...45R, 2013ApJ...777L..31D}.

To further elucidate this subject, we use the tree based N-body code {\sc ChaNGa}\citep{2008IEEEpds...ChaNGa, 2015AphCom..2..1}  to follow the 
collisional evolution of a planetesimal disc under the gravitational influence of a Jupiter-sized body. Because particle positions are sorted into a tree 
structure, neighbor finding and collision detection can be done quickly and efficiently. This considerably relaxes the constraints on resolution and 
integration time. With this toolset, we explore the collision rate structure of a planetesimal disc in the vicinity of mean-motion resonances with a planet. In 
particular, we would like to determine (1) what dynamics govern the collision rate profile near MMRs, (2) whether MMRs leave a detectable signature in the
collisionally-generated dust, and (3) whether these signatures can be used to determine or constrain the orbital properties of the perturbing planet.

This work is organized in the following way: In section \ref{sec:dynamics}, we provide an overview of the relevant dynamics that drive the evolution 
of a planetesimal disc under the gravitational influence of an external perturber. In section \ref{sec:sims} we provide an overview of the N-body code 
used and describe the initial conditions chosen for five simulations in which a perturbing giant planet is given various masses  and eccentricities. Section
\ref{sec:results} presents the results of these simulations and we take an in-depth look at the collision rate profiles of the planetesimals near the MMRs. 
In section \ref{sec:dust},  we extrapolate the collision rate down to smaller bodies and motivate a direct correspondence between the planetesimal collision profile
and the resulting dust profile. In section \ref{sec:constrain}, we use the resolved collisions to generate synthetic dust 
emission profiles that would be detected with observing facilities like ALMA. Under our simplifying assumptions, we show that a characteristic bump 
or dip feature appears in the dust emission near the interior 2:1 MMR, the presence of which depends on the mass and eccentricity of the perturbing 
planet. If the 2:1 MMR can be identified (presumably, by identifying another prominent MMR, such as the 3:1, and measuring the spacing between the two), we discuss the 
potential for using this feature to place constraints on the mass and eccentricity of the planet. In addition, we highlight the caveats of connecting the dust and planetesimal collision 
profiles in such a simple way. Finally, we conclude in section \ref{sec:conclusions}.

@@

We begin by providing a description of the dynamical effects responsible for shaping the orbital distribution of the planetesimals. The purpose of this is twofold: (1) to motivate
the initial conditions used for the simulations described in section \ref{sec:ics} and (2) to justify the exclusion of certain physical effects from our simulations. Here, we focus solely on the 
physics relevant for full-sized ($\sim$100 km) planeteismals and save a discussion of the effects on smaller bodies generated through collisions for section \ref{sec:dust}.

@@

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and a planetesimal, 
respectively. $e_{g}$ and $\varpi_{g}$ are the eccentricity and
longitude of pericenter of the giant planet, and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{1999ssd..book.....M} (ch. 6, pg. 237, eq. 6.67) as

@@

For the case of a Jupiter mass planet at 5.2 au perturbing a test particle at 3 au, the timescale for secular forcing is approximately 12,000 years. The details of this calculation
can be found in appendix \ref{sec:sec_forcing_timescale}.

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a
body, with $M$ being the mean anomaly. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Furthermore, variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citep{1999ssd..book.....M}. An important 
point to note, which we will revisit later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of $\phi$. 
Calculations for the widths of first and second order interior MMRs are shown in appendix \ref{sec:libration}, along with a timescale for the libration. For the most prominent interior 
mean-motion resonances, including the 2:1 and 3:1, the timescale associated with these oscillations driven by a Jupiter mass planet is  $\sim$ 1,000 - 2,000 years.

@@

A simple analytic model for the collision rate of a planetesimal population is given by \citet{1969edo..book.....S} as

@@

\noindent where $n$ is the number density of the population, $s$ and $m$ are the radii and masses of the bodies and $v$ is their typical encounter velocity. The encounter velocity is often described in terms of the rms eccentricity $\left<e^{2}\right>^{1/2}$ and inclination $\left<i^{2}\right>^{1/2}$ of the population (e.g. \citet{1993ARA&A..31..129L}) as

@@

\noindent where $v_{k}$ is the local Keplerian velocity. The second term in equation \ref{eq:saf} can be thought of as an additional enhancement to the collision cross section due to 
gravitational focusing. When the typical encounter velocity is small compared to the mutual escape velocity of the planetesimals, the collision cross section greatly exceeds the geometric 
value. An important feature of equation \ref{eq:saf} is that for a fixed value of $n$, the collision rate exhibits a global minimum as a function of $v$. For small $v$, gravitational focusing 
facilitates more collisions, while for large $v$ the encounter rate simply becomes so great that the collision rate again increases, even though gravitational focusing is mostly suppressed. 
This point will become relevant in section \ref{sec:vary_ecc} when we examine the qualitative changes in the collision rate near mean-motion resonances.

Although equation \ref{eq:saf} works well to describe the collision rate for a homogeneous collection of planetesimals, some problems arise when regions of commensurabilities are 
introduced. Namely, the resonances cause the orbits of planetesimals to precess, and an interface between secularly aligned and randomly oriented orbits arise on each side of the 
resonance. Additionally, an interface between dynamically cold and hot planetesimals develop. These two effects cause the number density, collision cross section and encounter velocity 
to rather abruptly vary across the boundaries of the resonance. For this reason, an N-body treatment in which collisions are directly resolved is necessary to understand how the collision 
rate varies near the MMRs.

@@

Over the course of many orbits, the residual gas from the primordial nebula can damp the eccentricities and inclinations of bodies. For a planetesimal-sized body, gas drag operates 
in the Stokes regime and the timescale for aerodynamic forces to significantly alter its relative velocity is given by \citep{1976PThPh..56.1756A}

@@

\noindent where $m$ and $s$ are the mass and radius of the planetesimal. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{g}$ is the headwind velocity of the gas experienced by the planetesimal. At 3 au, the gaseous component of the solar nebula has a
density of $4 \times 10^{-10}$ g cm$^{-3}$ and the typical headwind experienced by a body on a Keplerian orbit is $\sim 5,000$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 100 km 
body with a density of 2 g cm$^{-3}$ (assuming a mixture of ice and rock) the stopping timescale is on the order of 1 Myr. This is much longer than the timescales associated with 
secular forcing and libration due to mean-motion resonances, as discussed above, but shorter than the typical protoplanetary disk lifetime and potentially shorter the planetesimal 
formation timescale. For this reason, we do not model the effects of gas drag in the simulations, although we consider its effects when constructing initial conditions, which is discussed in 
the next section.

@@

To follow the dynamical and collisional evolution of a planetesimal disc, we use the highly parallel N-body code {\sc ChaNGa} 
\footnote{A public version of {\sc ChaNGa} can be downloaded from \url{http://www-hpcc.astro.washington.edu/tools/ChaNGa.html}}. 
This code, which is written in the {\sc CHARM++} parallel programming language, was originally designed for cosmology simulations 
and has been shown to perform well on up to half a million processors \citep{2015AphCom..2..1}. Using similar methods to {\sc PKDGRAV}, which has been used for numerous 
studies of planet formation \citep{2000Icar..143...45R, 2005ApJ...625..427L, 2013ApJ...777L..31D, 2015ApJ...806...23L}, {\sc ChaNGa} calculates 
gravitational forces using a modified Barnes-Hut \citep{1986Natur.324..446B} tree with hexadecapole expansions of the moments and integrates the equations of motion using a 
kick-drift-kick leapfrog scheme. All of the simulations we perform use a node opening criterion of $\theta_{BH}$ = 0.7. More information about the code can be found in 
\citet{2008IEEEpds...ChaNGa}.

We have recently modified {\sc ChaNGa} to handle solid-body collisions between particles, by assigning them a fixed radius, rather than treating them as tracers of a fluid with a 
characteristic softening length. We provide a brief summary of the collision model here, although a full description is provided in \citet{2019MNRAS.489.2159W}. This work is largely based 
on the solid-body collision implementation in {\sc PKDGRAV}, which is detailed in \citet{1994MNRAS.269..493R} and \citet{2000Icar..143...45R}. Imminent collisions are found during the 
drift phase of each time step by extrapolating the positions of bodies forward using the velocity calculated during the first kick. For each body, the nearest 64 neighbors are tested for an 
imminent collision. If an imminent collision is detected, the two particles of mass $m_{1}$ and $m_{2}$ are merged together together to form one single larger body with the same density 
and a mass of $m_{1} + m_{2}$. The resulting body is then imparted
with the center of mass position and velocity of the two
colliders. Because the resolution of a collision can result in 
another imminent collision, these events must be handled one at a time, with the soonest collision being resolved first. For this reason, another collision search is run each time a collision 
is resolved, which continues until there are no more imminent collisions during the current time step. During the course of the simulation, no debris is created by collisions. To model the 
collisionally generated dust distribution, we use the statistics of
the resolved planetesimal collisions to build a dust profile. This process is discussed in detail in section \ref{sec:dust}.

@@

In total, five simulations are run, which are listed in table \ref{tab:sims}. The second is a ``nominal'' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass or eccentricity is altered by a factor of two from the nominal value. In all cases, the perturbing giant is placed on a 5.2 au orbit around a 1 
$M_{\odot}$ star. The planetesimal disc extends from 2.2 to 3.8 au, which covers the two most prominent mean-motion resonances with the giant planet, the 2:1 at 3.27 au and the 3:1 at 
2.5 au. The planetesimal disc follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

\noindent with $\alpha$ = 3/2 and $\Sigma_{0}$ = 10 g cm$^{-2}$. Planetesimals are given a bulk density of 2 g cm$^{-3}$, which corresponds to a mixture of ice and rock, and 
are given a diameter of 300 km. This choice of parameters is meant to mimic the setup used in simulation B from \citet{2000Icar..143...45R}. One difference from the 
aformentioned study is that we use a narrower annulus, which reduces the number of particles required from $10^6$ down to roughly 500,000. This allows us to use a finer base timestep 
size, which, as we will discuss in section \ref{sec:results}, appears to be the reason why no features were seen in the collision profile near the resonances by \citet{2000Icar..143...45R}.

Another difference from \citet{2000Icar..143...45R} is that the dynamical effects of secular forcing by the giant planet, along with the effects of viscous stirring and gas drag on the 
planetesimals, are built into the initial conditions. Although we do not model the effects of gas drag on the planetesimals during the simulation, we construct the initial conditions 
such that the effects of gas drag and viscous stirring are in balance. The viscous stirring timescale of the planetesimal disc is much longer than our chosen integration time, and so the 
dynamical excitation of the disc (excluding resonances) stays constant, even without the inclusion of damping forces from the gas. This is done by first calculating the equilibrium 
eccentricity $e_{eq}$ due to viscous stirring and gas drag as a function of semimajor axis according to equation 12 of \citet{2002ApJ...581..666K}. The eccentricities of the bodies are 
drawn randomly from a Rayleigh distribution with a mode of $e_{eq}$, while the inclinations are drawn from a similar distribution with a mode of $e_{eq}$/2 \citep{1993MNRAS.263..875I}. 
The arguments of perihelion $\omega$, longitude of ascending nodes $\Omega$ and the mean anomalies $M$ of the bodies are drawn uniformly $\in [0, 2 \pi)$.

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. Through experimentation, we have found that using a base timestep size with {\sc ChaNGa} of 3\% of an orbital dynamical time
at the inner edge of the disc keeps the integration symplectic. Although doing so preserves orbital frequencies, the errors associated with
precession frequencies do not average out to zero. This is especially important given the effects of secular forcing by the planet. To mitigate
this, we further reduce the base timestep size by a factor of 4 and use $\Delta t$ = 0.0025 yr. Doing so prevents the longitude of perihelia of
planetesimals at the inner edge of the disc, where this artificial precession is most severe, from drifting by more than the intrinsic spread in $\varpi$ due to the
free eccentricity over the course of our integrations.

An additional timescale is set by the dynamical time of the planetesimals ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. This timescale must be resolved in order to properly 
follow close gravitational encounters between these bodies. To resolve the base time step and the dynamical timescale of planetesimals simultaneously, we use a two-tiered time 
stepping scheme, following\citep{2015ApJ...806...23L}. To start, all bodies are placed on the base time step. A first pass of collision detection is then run in which the radii of all bodies are 
inflated by a factor of 2.5. Any bodies with imminent collisions predicted using the inflated radii are placed on a time step that is a factor of 16 smaller than the orbital time step. Although 
this is still a factor of roughly 7 larger than the dynamical time of a planetesimal, we found no difference in the collision rate when using any smaller of a minimum time step size.The 
purpose of the two-tiered scheme is to properly resolve the gravitational interactions between any bodies that undergo a close encounter. This prevents the coarser base time step from 
reducing the effectiveness of gravitational focusing, while minimizing the additional computational expense.
